---
title: "Regression-Models-Project"
author: "Abhijeet"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Synopsis
The following analysis is done on the "mtcars" data in the R dataset package and addresses the following questions:

1. “Is an automatic or manual transmission better for MPG”

2. "Quantify the MPG difference between automatic and manual transmissions" 
- Description: The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

## Data Processing
```{r}
#libraries
library(datasets);library(ggplot2);require(stats); require(graphics)
data(mtcars)
```
## Exploratory Data Analysis
```{r}
?mtcars
summary(mtcars)
```
The data frame contains 32 observations on 11 (numeric) variables.

1. mpg	Miles/(US) gallon
2. cyl	Number of cylinders
3. disp	Displacement (cu.in.)
4. hp	Gross horsepower
5. drat	Rear axle ratio
6. wt	Weight (1000 lbs)
7. qsec	1/4 mile time
8. vs	Engine (0 = V-shaped, 1 = straight)
9. am	Transmission (0 = automatic, 1 = manual)
10. gear	Number of forward gears
11. carb	Number of carburetors
```{r}
boxplot(mtcars$mpg~mtcars$am,main="mtcars data",xlab="mpg(Miles/(US) gallon)",ylab="am(Transmission)")
```
Checking Normality
```{r}
shapiro.test(mtcars$mpg)
```
- The P-value of the Shapiro-Wilk normality test must be greater than 0.05 for assuming normality. Hence, the mpg(Miles/(US) gallon
) data is normally distributed. 
```{r}
t.test(mtcars$mpg~mtcars$am, paired = FALSE, var.equal = FALSE)
```
- From the above test it is ensured that null Hypothesis(p-value < Significance level) can be rejected at 5% significance level. The mpg(miles per gallon) is higher for manual transmission and there is a significant difference between the auto and manual transmission.

## Regression model

### logistic regression

First, lets check the dependency of miles per gallon(mpg) and transmission(am), particularly looking to answer the following  question:

“Is an automatic or manual transmission better for MPG?"

In this test the outcome is binary or categorical variable 0 and 1.Thus use logistic regression.
```{r}
fit<-glm(am~mpg,data=mtcars, family ="binomial")
summary(fit)
exp(fit$coefficients)
```
AIC which similar to R² tells the significance of test but opposite of R² the lesser the value of AIC better fit to the data. In this test the AIC is 33.675.There is a 36% probability of a transmission to be manual for every additional mile per gallon. To improve AIC, trying multivariate regression.

### multivariate regression
The best fit model suggested that the transmission, weight and 1/4 mile time/ performance measure of acceleration are the best fir variables.
```{r}
fit_1<-lm(formula = mpg ~ wt + qsec + am, data = mtcars)
summary(fit_1)$coef
summary(fit_1)$r.squared
```
## Conclusion
In logistic regression it can observed that there is a 36% probability of a transmission to be manual for every additional mile per gallon. While it is difficult to interpret in multivariate regression because at significance level alpha=0.05, miles per gallon and transmission are significantly influenced by other factors( weight and acceleration).





_________________________________________________________________

## Appendix

### Appendix A -Data Visualization
```{r}
library(GGally)
options(repr.plot.width = 30, repr.plot.height = 30)
ggpairs(mtcars,  title = "Scatter and correlation matrix",upper = list(continuous = wrap("cor",size = 3)), 
    lower = list(continuous = wrap("smooth")))
```
Normality visualization
```{r}
qqnorm(mtcars$mpg)
qqline(mtcars$mpg,col=2)
```

### Appendix B - Model Selection
```{r}
step(lm(mpg~.,data=mtcars),direction = "both",trace=0)
```
### Appendix C logistic regression Plots
Plot of residuals for my multivariate regression.

```{r}
par(mfrow = c(2, 2))
plot(fit)
```

### Appendix D Multivariate Regression
```{r}
par(mfrow = c(2, 2))
plot(fit_1)
```
